---
title: "writing_functions"
author: "Mukta Patwari"
date: "2025-10-23"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Start small

z-scores

```{r}
x_vec = rnorm(20, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Function to compute z scores

```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
    stop("Only compute z scores when the input has 5 or more numbers")
  }
  
  z = (x - mean(x)) / sd(x)
  
  z
}
```

Let's try our function...

```{r}
z_scores(x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)
```

Let's break our function

```{r, error = TRUE}
z_scores(3)

z_scores("my name is Mukta")
```

## Computing

Compute and return mean and SD of a numeric vector

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  if (length(x) < 5) {
    stop("Only compute z scores when the input has 5 or more numbers")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}

mean_and_sd(x_vec)
```

Simulation

```{r}
sim_df =
  tibble(x = rnorm(n = 30, mean = 3, sd = 2))

sim_df %>% 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
```

Write a function to do simulations

(We wrote this code chunk last time, now it's being sourced.)

```{r}
source("source/sim_mean_sd.R")
```

Let's run this function. This is providing a named argument.

```{r}
sim_mean_and_sd(n_subj = 50)
```

What if you just put "50"? By default, it assigns the value to the first argument.

```{r}
sim_mean_and_sd(50)
```

Import LoTR data

```{r}
fellowship_ring =
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "Fellowship of the Ring")

two_towers =
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "Two Towers")

return_of_the_king =
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "Return of the King")

lotr_df =
  bind_rows(fellowship_ring, two_towers, return_of_the_king) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words")  %>%  
  mutate(race = str_to_lower(race)) %>%  
  select(movie, everything()) 
```

Turn this into a function

```{r}
lotr_import = function(cell_range, movie_title) {
  
  df =
    read_excel("data/LotR_Words.xlsx", range = cell_range) %>%
    mutate(movie = movie_title)
  
  df
  
}

fellowship = lotr_import(cell_range = "B3:D6", movie_title = "Fellowship")
two_tower = lotr_import(cell_range = "F3:H6", movie_title = "Two Towers")
return = lotr_import(cell_range = "J3:L6", movie_title = "Return of the King")

bind_rows(fellowship, two_towers, return)
```

Look at one more example

```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}
```

